---
title: "Extending EDI with Custom R6 Inference Classes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extending EDI with Custom R6 Inference Classes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

EDI is implemented with R6 classes. Advanced users can define their own R6
classes outside the package and reuse EDI's design storage, response handling,
randomization, bootstrap, and summary methods.

The custom-extension base classes are intentionally internal while the extension
contract is experimental. Retrieve them with `getFromNamespace()`:

```{r}
library(EDI)
library(R6)

InferenceCustomAsymp <- getFromNamespace("InferenceCustomAsymp", "EDI")
```

## Inference Contract

A custom asymptotic inference class should inherit from `InferenceCustomAsymp`
and implement a public `fit(estimate_only = FALSE)` method. `fit()` returns a
named list with:

- `estimate`: required numeric scalar treatment-effect estimate.
- `se`: optional numeric scalar standard error.
- `df`: optional degrees of freedom. Use `NA_real_` for z inference.
- `model`: optional fitted model object retained by `get_mod()`.
- `nonestimable_reason`: optional character scalar used when the estimate or
  standard error is unavailable.

Use public accessors instead of private fields:

- `get_response()`
- `get_treatment()`
- `get_covariates()`
- `get_analysis_data()`
- `get_design_object()`
- `get_response_type()`

## Example

```{r}
InferenceMedianDiff <- R6Class(
  "InferenceMedianDiff",
  inherit = InferenceCustomAsymp,
  public = list(
    fit = function(estimate_only = FALSE) {
      dat <- self$get_analysis_data()
      y_t <- dat$y[dat$w == 1]
      y_c <- dat$y[dat$w == 0]

      est <- stats::median(y_t) - stats::median(y_c)
      if (estimate_only) {
        return(list(estimate = est))
      }

      list(
        estimate = est,
        se = sqrt(stats::var(y_t) / length(y_t) + stats::var(y_c) / length(y_c)),
        df = length(y_t) + length(y_c) - 2,
        model = NULL
      )
    }
  )
)

des <- FixedDesignBernoulli$new(n = 20, response_type = "continuous")
des$add_all_subjects_to_experiment(data.frame(x = seq_len(20)))
des$overwrite_all_subject_assignments(rep(c(0, 1), each = 10))
des$add_all_subject_responses(rnorm(20))

inf <- InferenceMedianDiff$new(des)
inf$compute_estimate()
inf$compute_asymp_two_sided_pval()
inf$compute_asymp_confidence_interval()
inf$compute_bootstrap_two_sided_pval(B = 501)
```

## Custom Designs

EDI also provides internal bases for user-defined designs:

```{r}
DesignCustomFixed <- getFromNamespace("DesignCustomFixed", "EDI")
DesignCustomSequential <- getFromNamespace("DesignCustomSequential", "EDI")
```

For fixed designs, implement `draw_assignments(r)` and return an `n x r` 0/1
assignment matrix. For sequential designs, implement `assignment_rule()` and
return a scalar 0/1 assignment for the current subject. EDI handles subject
storage, response recording, and validation.
